|

### 2025-10-29 output/Response_to_action_editor_reviewers-2.Rmd

(Please note also the update in the title page to acknowledge the colleague who brought this previous literature to our attention, as well as the contributions of the reviewers.)

# Responses to Reviewer 4:

###### It is always best, many would say required, to label the axes. The lack of axis labels in the graphics of Figure 1 make the rest of these comments difficult to type, so for convenience I imagine the axes are labeled conventionally x and y in the left panel, and x and y' in the right panel, indicating that y' is a rotated representation of y. (In the ms., it would be best to tie the labeling notation to the algebra used.)

We had overlooked labeling the axes and have done so in this new version;
we thank the reviewer once again for pointing this out.
$`r TRAIT_COMP_1`$/$`r TRAIT_COMP_2`$ have been used as labels for simplicity
(both in the code for plotting them, and for the reader's understanding),
but note that they represent the coordinates of
a latent trait point transformed to the test space,
that is, $`r TRAIT_COORD_1`$/$`r TRAIT_COORD_2`$.
However, there is a direct correspondence between the algebraic components and
the coordinates in either space, so this labeling should be clear enough,
while relying only upon the algebraic interpretation of the model,
thus avoiding the geometrical details,
which are only necessary for deriving the results.
This choice of labeling is explained in the text in p. 23, ll. 1-8 (see also our response to the last comment below):

> In formal terms, the proper axis labeling of this figure
> should correspond to the vectors that make up basis $`r TEST_SPACE_BASIS`$,
> in this case, $`r ITEM_BASIS_VECTOR_FIRST`$ and $`r ITEM_BASIS_VECTOR_2`$;
> the coordinates represent the coefficients that multiply each basis vector to
> obtain the corresponding coordinate in the test space.
> However, these coefficients correspond to the values of
> a vector in the test space, be it $`r DISCR_COORDS`$ for $`r DISCR_VECTOR`$,
> or $`r TRAIT_COORDS_TRANSF_TS`$ for $`r TRAIT_VECTOR`$
> (as given by Equation \@ref(eq:trait-coords-spaces-transform)).
> For the sake of simplicity,
> we label the axes as $`r TRAIT_COMP_1`$ and $`r TRAIT_COMP_2`$,
> highlighting the comparison of the items with the latent trait vectors.
> However, note that these coordinates actually correspond to
> the components of $`r TRAIT_COORDS_TRANSF_TS`$, i.e.,
> $`r TRAIT_COORD_1`$ and  $`r TRAIT_COORD_2`$.

@@

> 
> A comparison of panels (a) and (c) allows understanding the effect of
> a positive correlation on the $`r MDISC_CORR_PARAM`$
> (and consequently the $`r MIL_CORR_PARAM`$) value of an item: [...]

Finally, we have added a fourth panel (panel (d)) that illustrates how the items would be plotted if we omitted using the proper test space basis (and just using the canonical basis instead).
Incidentally, this also serves the purpose of allowing an easier comparison of the coordinates computed from the two versions (the agnostic, and the correlation-based one) of the multidimensional parameters.
We have briefly explained this in p. 24, ll. 17-25 (please refer to the changes explained above).

### 2025-11-04 output/Response_to_action_editor_reviewers-2.Rmd

# Responses to Reviewer 4:

###### It is always best, many would say required, to label the axes. The lack of axis labels in the graphics of Figure 1 make the rest of these comments difficult to type, so for convenience I imagine the axes are labeled conventionally x and y in the left panel, and x and y' in the right panel, indicating that y' is a rotated representation of y. (In the ms., it would be best to tie the labeling notation to the algebra used.)

We had overlooked labeling the axes and have done so in this new version;
we thank the reviewer once again for pointing this out.
$`r TRAIT_COMP_1`$/$`r TRAIT_COMP_2`$ have been used as labels for simplicity
(both in the code for plotting them, and for the reader's understanding),
but note that they represent the coordinates of
a latent trait point transformed to the test space,
that is, $`r TRAIT_COORD_1`$/$`r TRAIT_COORD_2`$.
However, there is a direct correspondence between the algebraic components and
the coordinates in either space, so this labeling should be clear enough,
while relying only upon the algebraic interpretation of the model,
thus avoiding the geometrical details,
which are only necessary for deriving the results.
This choice of labeling is explained in the text in p. 23, ll. 1-8 (see also our response to the last comment below):

> In formal terms, the proper axis labeling of this figure
> should correspond to the vectors that make up basis $`r TEST_SPACE_BASIS`$,
> in this case, $`r ITEM_BASIS_VECTOR_FIRST`$ and $`r ITEM_BASIS_VECTOR_2`$;
> the coordinates represent the coefficients that multiply each basis vector to
> obtain the corresponding coordinate in the test space.
> However, these coefficients correspond to the values of
> a vector in the test space, be it $`r DISCR_COORDS`$ for $`r DISCR_VECTOR`$,
> or $`r TRAIT_COORDS_TRANSF_TS`$ for $`r TRAIT_VECTOR`$
> (as given by Equation (24)).
> For the sake of simplicity,
> we label the axes as $`r TRAIT_COMP_1`$ and $`r TRAIT_COMP_2`$,
> highlighting the comparison of the items with the latent trait vectors.
> However, note that these coordinates actually correspond to
> the components of $`r TRAIT_COORDS_TRANSF_TS`$, i.e.,
> $`r TRAIT_COORD_1`$ and  $`r TRAIT_COORD_2`$.

### 2025-11-04 output/IMPS-2025-slideshow.qmd

::: {.column .fragment}
```{r item-plot-out}
#| fig-cap:   Item representation in non-orthonormal test space
#| fig-align: center
transform_grid(
  ts_basis,
  x_limits   = c(-2, 3),
  y_limits   = c(-2, 2.8),
  break_step = 1
) +
  geom_segment(
    arrow     = arrow(angle = 20, length = unit(10, "points"), type = "closed"),
    linejoin  = "mitre",
    linewidth = VECTOR_WIDTH,
    data      = items_oblique |> arrange(desc(item)),
    mapping   = aes(
      origin_transf_1, origin_transf_2,
      xend  = end_transf_1, yend = end_transf_2,
      color = item, fill = item
    )
  ) +
  scale_color_manual(values = PALETTE, guide = NULL) +
  coord_fixed(expand = FALSE, clip = "on")
```
:::
:::::

>>>>

::: {.column .fragment}
```{r item-plot-out}
#| fig-cap:   Item representation in non-orthonormal test space
#| fig-align: center
transform_grid(
  ts_basis,
  x_limits   = c(-2, 3),
  y_limits   = c(-2, 2.8),
  break_step = 1
) +
  items_geom_oblique +
  colorscale_no_legend +
  coord_fixed(expand = FALSE, clip = "on")
```
:::
:::::

### 2025-11-04 output/Title_page.Rmd

[swh:1:dir:3914435e67a2b34847e36a712a181d7d9d76f35e;origin=https://github.com/DaniMori/mirt-parameters;visit=swh:1:snp:9069b256471ea08d5748e8bb021b7ee876aee724;anchor=swh:1:rev:fe49f438bf940606ae7e3c86c3c55f202e6920af](https://archive.softwareheritage.org/swh:1:dir:3914435e67a2b34847e36a712a181d7d9d76f35e;origin=https://github.com/DaniMori/mirt-parameters;visit=swh:1:snp:9069b256471ea08d5748e8bb021b7ee876aee724;anchor=swh:1:rev:fe49f438bf940606ae7e3c86c3c55f202e6920af)

>>>>

[swh:1:dir:6195936c6e16023c4a0b6a4500b81a34469d64f1;origin=https://github.com/DaniMori/mirt-parameters;visit=swh:1:snp:410cf1a6fb4dd4c10bdf0ce7f12f6890f560549d;anchor=swh:1:rev:1f1a708a7477dd61c0ba5c82dcf1d8523ab3d8ff](https://archive.softwareheritage.org/swh:1:dir:6195936c6e16023c4a0b6a4500b81a34469d64f1;origin=https://github.com/DaniMori/mirt-parameters;visit=swh:1:snp:410cf1a6fb4dd4c10bdf0ce7f12f6890f560549d;anchor=swh:1:rev:1f1a708a7477dd61c0ba5c82dcf1d8523ab3d8ff)

### 2025-11-04 output/seminar-slideshow.qmd

plot_orth <- plot_orth +
  scale_color_manual(values = ITEMS_PALETTE) +
  theme(
    axis.ticks       = element_blank(),
    legend.title     = element_blank(),
    panel.grid.minor = element_blank()
  )

plot_oblique <- plot_oblique +
  scale_color_manual(values = ITEMS_PALETTE, guide = NULL) +
  theme(axis.ticks = element_blank())

>>>>

plot_orth <- plot_orth_uncorr_contours +
  theme(
    axis.ticks       = element_blank(),
    legend.title     = element_blank(),
    panel.grid.minor = element_blank()
  )

plot_oblique <- plot_oblique_ts +
  theme(axis.ticks = element_blank())
